---
title: "Insights into temporal variability and reproducibility of scent profiles in Antarctic fur seals - R Script"
author: "Jonas Tebbe"
date: ""
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(GCalignR)
library(vegan)
library(readr)
library(ggplot2)
library(ggbeeswarm)
library(tidyverse)
```

## Alignment and preliminary data properties
```{r Alignment and preliminary data properties, message=FALSE, warning=FALSE}
## Load and view GCalignR alignment objects for GCMS scent data in two and six breeding beaches
load("RData/objects/mom_pup_alignment_GCalignR.RData")
mom_pup_aligned

load("RData/objects/pup_colonies_alignment_GCalignR.RData")
pup_colonies_aligned

## Load raw information for all samples containing raw peaks and calculate mean peak number
load("RData/objects/seal_raw_dfs.Rdata")

individual_peak_number <- NULL
for (i in 1:length(seal_dfs.list)) {
  individual_peak_number[i] <- length(seal_dfs.list[[i]]$RT)
}

mean_ind_peaks <- mean(individual_peak_number)
sd_ind_peaks <- sd(individual_peak_number)
cat("\n", "\n", "Mean peaks:", as.character(mean_ind_peaks), "\n", "Peak SD:", as.character(sd_ind_peaks))
```


## NMDS scaling of mother-pup alignment data
```{r NMDS scaling of mother-pup alignment data, eval=FALSE, include=TRUE}
load("RData/objects/mom_pup_alignment_GCalignR.RData")

scent_factors_raw <- read_delim("documents/metadata_seal_scent.txt", 
                                "\t", escape_double = FALSE, trim_ws = TRUE)
scent_factors_raw <- as.data.frame(scent_factors_raw[-c(194:209),])

# set sample names as row names, ensure there are no duplicates
scent_factors <- scent_factors_raw[,-1]
rownames(scent_factors) <- scent_factors_raw[,1]

## check for empty samples, i.e. no peaks
x <- apply(mom_pup_aligned$aligned$RT, 2, sum)
x <- which(x == 0)

## normalise area and return a data frame
scent <- norm_peaks(mom_pup_aligned, conc_col_name = "Area",rt_col_name = "RT",out = "data.frame") 
## common transformation for abundance data to reduce the extent of mean-variance trends
scent <- log(scent + 1) 

## subset scent_factors
scent_factors <- scent_factors[rownames(scent_factors) %in% rownames(scent),]
scent <- scent[rownames(scent) %in% rownames(scent_factors),]

## keep order of rows consistent
scent <- scent[match(rownames(scent_factors),rownames(scent)),] 

## get number of compounds for each individual sample after alignment
num_comp <- as.vector(apply(scent, 1, function(x) length(x[x>0]))) 

## bray-curtis similarity
scent_nmds.obj <- vegan::metaMDS(comm = scent, k = 2, try = 999, trymax = 9999, distance = "bray") 

scent_nmds <- as.data.frame(scent_nmds.obj[["points"]])  

scent_nmds <- cbind(scent_nmds,
                    age = scent_factors[["age"]],
                    tissue_tag = scent_factors[["tissue_tag"]],
                    colony = scent_factors[["colony"]],
                    family = as.factor(scent_factors[["family"]]),
                    clr = as.factor(scent_factors[["clr"]]),
                    shp = as.factor(scent_factors[["shp"]]),
                    gcms = as.factor(scent_factors[["gcms_run"]]),
                    peak_res = as.factor(scent_factors[["peak_res"]]),
                    sample_qlty = as.factor(scent_factors[["sample_qlty"]]),
                    vialdate = as.factor(scent_factors[["gcms_vialdate"]]),
                    captured = as.factor(scent_factors[["capture_date"]]),
                    sex = scent_factors[["sex"]],
                    num_comp = num_comp)
scent_nmds <- scent_nmds %>% mutate(BeachAge = str_c(colony, age, sep = "_")) # creates & adds new variable BeachAge Colony Identity + Age(Mom or Pup) of the individual
```


## Colony membership in SSB and FWB mom-pup pairs
```{r Colony membership in SSB and FWB mom-pup pairs, echo=TRUE, message=FALSE, warning=FALSE}
load("RData/objects/mom_pup_nmds_scaling.RData")

mp_colony_gg <- ggplot(data = scent_nmds) + 
  geom_point(size = 4.5, aes(MDS1, MDS2, color = BeachAge, shape = BeachAge)) + 
  scale_shape_manual(values = c(19, 1, 19, 1), 
                     labels = c("FWB mothers ", "FWB pups ", "SSB mothers ", "SSB pups ")) +
  scale_color_manual(values = c("#D55E00", "#D55E00", "#56B4E9", "#56B4E9"), 
                     labels = c("FWB mothers ", "FWB pups ", "SSB mothers ", "SSB pups ")) +
  theme_void() + 
  ylim(-0.75,1.1) +
  # ggtitle("Colony Differences SSB & FWB '16/'17") +
  annotate("text", x = 0.64, y = 1.1, label = "(A)", size = 5) +
  annotate("text", x = 0.48, y = -0.75, label = "2D Stress: 0.23", size = 5) +
  theme(panel.background = element_rect(colour = "black", size = 1, fill = NA),
        aspect.ratio = 1,
        legend.position = "none",
        legend.title = element_blank(),
        legend.background = element_rect(size = 0.3, linetype = "solid", color = "black")) 
mp_colony_gg
```


## Family ID in SSB & FWB mother-pup pairs
```{r Family ID in SSB & FWB mother-pup pairs, echo=TRUE, message=FALSE, warning=FALSE}
load("RData/objects/mom_pup_nmds_scaling.RData")

# create color palette for the plot
clr <- c("#D55E00", "red", "#56B4E9", "#009E73","#000000", "#CC79A7") 

# assign pch values for plotting
shp <- c(0,1,2,7,10,5,6,18,16,17,15) 

# create unique color-pch pairs
color_shape_pairs <- crossing(clr,shp) 

# randomly sample 50 unique pairs (sample without replacement)
set.seed(123) # always get same pairs in a run
color_shape_pairs <- color_shape_pairs[sample(nrow(color_shape_pairs), 50),] 

# assign new dataframes to transform scent_nmds$clr & shp with the unique values we created
color_shape_pairs_plot <- rbind(color_shape_pairs[1:25,],color_shape_pairs[1:7,],color_shape_pairs[7,],  color_shape_pairs[8:25,], color_shape_pairs[26:50,], color_shape_pairs[26:50,])
scent_nmds$clr <- as.factor(color_shape_pairs_plot$clr)
scent_nmds$shp <- as.factor(color_shape_pairs_plot$shp)

# family plot season 16/17
mp_family_gg <- ggplot(data = scent_nmds,aes(MDS1,MDS2, color = clr, shape = shp)) + 
  geom_point(size = 4.5) +
  scale_shape_manual(values = as.numeric(levels(scent_nmds$shp))) +
  theme_void() + 
  ylim(-0.75,1.1) +
  scale_color_manual(values = levels(scent_nmds$clr)) +
  annotate("text", x = 0.64, y = 1.1, label = "(B)", size = 5) +
  annotate("text", x = 0.48, y = -0.75, label = "2D Stress: 0.23", size = 5) +
  theme(panel.background = element_rect(colour = "black", size = 1,
                                        fill = NA), aspect.ratio = 1, legend.position = "none") 
mp_family_gg
```


## NMDS scaling for alignment data in six pup colonies
```{r NMDS scaling for alignment data in six pup colonies, eval=FALSE, include=TRUE}
load("RData/objects/pup_colonies_alignment_GCalignR.RData")
scent_factors_raw <- read_delim("documents/metadata_seal_scent.txt", 
                                "\t", escape_double = FALSE, trim_ws = TRUE)
scent_factors_raw <- as.data.frame(scent_factors_raw[-c(194:209),])

# set sample names as row names, ensure there are no duplicates
scent_factors <- scent_factors_raw[,-1]
rownames(scent_factors) <- scent_factors_raw[,1]

## check for empty samples, i.e. no peaks
x <- apply(pup_colonies_aligned$aligned$RT, 2, sum)
x <- which(x == 0)

## normalise area and return a data frame
scent <- norm_peaks(pup_colonies_aligned, conc_col_name = "Area",rt_col_name = "RT",out = "data.frame") 
## common transformation for abundance data to reduce the extent of mean-variance trends
scent <- log(scent + 1) 

## subset scent_factors
scent_factors <- scent_factors[rownames(scent_factors) %in% rownames(scent),]
scent <- scent[rownames(scent) %in% rownames(scent_factors),]

## keep order of rows consistent
scent <- scent[match(rownames(scent_factors),rownames(scent)),] 

## get number of compounds for each individual sample after alignment
num_comp <- as.vector(apply(scent, 1, function(x) length(x[x>0])))

## bray-curtis similarity
scent_nmds.obj <- metaMDS(comm = scent, k = 2, try = 999, trymax = 9999, distance = "bray") 
## MDS outcome evaluated with PCA for factor colony in metadata table for individuals
scent_nmds <- with(scent_factors, MDSrotate(scent_nmds.obj, colony))

## get x and y coordinates
scent_nmds <- as.data.frame(scent_nmds[["points"]])  

## add the colony as a factor to each sample
scent_nmds <- cbind(scent_nmds,
                    age = scent_factors[["age"]],
                    tissue_tag = scent_factors[["tissue_tag"]],
                    colony = scent_factors[["colony"]],
                    family = as.factor(scent_factors[["family"]]),
                    clr = as.factor(scent_factors[["clr"]]),
                    shp = as.factor(scent_factors[["shp"]]),
                    gcms = as.factor(scent_factors[["gcms_run"]]),
                    peak_res = as.factor(scent_factors[["peak_res"]]),
                    sample_qlty = as.factor(scent_factors[["sample_qlty"]]),
                    vialdate = as.factor(scent_factors[["gcms_vialdate"]]),
                    captured = as.factor(scent_factors[["capture_date"]]),
                    sex = scent_factors[["sex"]],
                    num_comp = num_comp)
# creates & adds new variable BeachAge Colony Identity + Age(Mom or Pup) of the individual
scent_nmds <- scent_nmds %>% mutate(BeachAge = str_c(colony, age, sep = "_")) 
```


## PERMANOVA for mother-pup pair comparison
```{r PERMANOVA for mother-pup pair comparison, echo=TRUE}
load("RData/objects/mom_pup_nmds_scaling.RData")

# set seed to reproduce the same outcome (can vary due to different permutations!)
set.seed(123)
adonis(scent ~ age+colony+colony:family, 
       data = scent_factors,
       method = "bray",
       permutations = 99999)

# test for group dispersal
mod <- betadisper(vegdist(scent), scent_factors$colony, type = "median")
anova(mod)

#Tukey Test
HSD.mod <- TukeyHSD(mod)
plot(HSD.mod)
HSD.mod
```


## PERMANOVA for comparison of six pup colonies
```{r PERMANOVA for comparison of six pup colonies, echo=TRUE}
load("RData/objects/pup_colonies_nmds_scaling.RData")
set.seed(123)
adonis(scent ~ age+colony+colony:family, 
       data = scent_factors, 
       permutations = 99999) 

# pairwise PERMANOVA
pairwiseAdonis::pairwise.adonis(scent, scent_factors$colony, perm = 99999)

# test for group dispersal
mod2 <- betadisper(vegdist(scent), scent_factors$colony, type = "median")
anova(mod2)
```


## Re-evaluation of 2011 field season scent data
Non-metric multidimensional scaling
```{r eval=TRUE, include=TRUE, echo=FALSE}
## normalise area and return a data frame
# uses intrinsic data of GCalignR, as this is data used in Stoffel et al. 2015
scent <- norm_peaks(aligned_peak_data, conc_col_name = "area",rt_col_name = "time",out = "data.frame") 
## common transformation for abundance data to reduce the extent of mean-variance trends
scent <- log(scent + 1) 

## GCalignR contains factors for the chemical dataset
data("peak_factors") 
## keep order of rows consistent
scent <- scent[match(row.names(peak_factors),row.names(scent)),] 
## NMDS using Bray-Curtis dissimilarities
scent_nmds <- vegan::metaMDS(comm = scent, distance = "bray", try = 999, trymax = 9999)
## get x and y coordinates
scent_nmds <- as.data.frame(scent_nmds[["points"]])  
## add the colony, etc. as a factor to each sample
scent_nmds <- cbind(scent_nmds,
                    colony = peak_factors[["colony"]],
                    family = peak_factors[["family"]],
                    age = peak_factors[["age"]])
scent_nmds <- scent_nmds %>% mutate(BeachAge = str_c(colony, age, sep = "_")) 
# creates & adds new variable BeachAge Colony Identity + Age(Mom or Pup) of the individual

#sort scent_nmds for color_shape_pairs_plot to form and plot unique family pairs
scent_nmds <- scent_nmds %>% arrange(colony)
```


Re-evalution in PERMANOVA instead of ANOSIM
```{r}
## PERMANOVA
set.seed(123)
adonis(scent ~ age+colony+colony:family, 
       data = peak_factors, 
       permutations = 99999)

# Test for heterogeneity
anova(betadisper(vegdist(scent), peak_factors$colony))
```


## Statistical analysis of NMDS data properties in all data frames
```{r Statistical analysis of NMDS data properties in all data frames, echo=TRUE, message=FALSE, warning=FALSE}
## compare number of compounds in moms and pups
load("RData/objects/mom_pup_nmds_scaling.RData")

mean(scent_nmds$num_comp[scent_nmds$age == "1"])

t.test(scent_nmds$num_comp[scent_nmds$age == "1"], 
       scent_nmds$num_comp[scent_nmds$age == "2"],
       paired = F)

## ANOVA for mom-pup differences based on the two different breeding beaches
mom.pup.aov <- aov(scent_nmds$num_comp ~ scent_nmds$BeachAge)
summary(mom.pup.aov)
mom.pup.aov

boxplot(scent_nmds$num_comp ~ scent_nmds$BeachAge,
        xlab = "Colony x Age", ylab = "Number of Compounds",
        frame = FALSE)

## compare number of compounds in all pup colonies with anova
  load("RData/objects/pup_colonies_nmds_scaling.RData")
pup.colony.aov <- aov(scent_nmds$num_comp ~ scent_nmds$colony)
summary(pup.colony.aov)
pup.colony.aov

boxplot(scent_nmds$num_comp ~ scent_nmds$colony,
        xlab = "Colony", ylab = "Number of Compounds",
        frame = FALSE)
```


## Effect size estimate by PERMANOVA R2 bootstrap
```{r Effect size estimate by PERMANOVA R2 bootstrap, eval=FALSE, include=TRUE}
## Load data and assign data to data.frames
load("RData/objects/R2_initial_season_btrap.RData")

old_season_colony <- paov_r2_results[[1]][[2]]
old_season_family <- paov_r2_results[[1]][[3]]

load("RData/objects/R2_replication_season_btrap.RData")
new_season_colony <- paov_r2_results[[1]][[2]]
new_season_family <- paov_r2_results[[1]][[3]]

MP_effectsize <- c(old_season_colony, new_season_colony, old_season_family, new_season_family)

MP_effectsize.groups <- c(rep("Colony S1", 5000),
                      rep("Colony S2", 5000),
                      rep("Family S1", 5000), 
                      rep("Family S2", 5000))

MP_effectsize.df <- data.frame(btrap_combined_results = MP_effectsize,
                           btrap_subset_groups = MP_effectsize.groups)
```


## Effect size estimate plot
```{r Effect size estimate plot, echo=TRUE}
load("RData/objects/effect_size_df.RData")
# point estimates for PERMANOVA on non-bootstrapped (original) data
point_estimate <- c(0.1444734, 0.09168289, 0.08780086, 0.1209394)
# point estimate groups for reasons of comprehensibility
point_estimate_groups <- c("Colony S1", "Colony S2", "Family S1", "Family S2")

# data is a dataframe with two columns: one with the values (adj_vals) and one which specifies to which species each value belongs to
MP_effectsize_gg <- ggplot(MP_effectsize.df, aes(y = btrap_combined_results, 
                                                 x = btrap_subset_groups, 
                                                 color = btrap_subset_groups)) + 
  # this arranges the points according to their density
  geom_quasirandom(alpha = 0.06, size = 3, width = 0.3, bandwidth = 1) + #width = 0.47, bandwidth = 2.5
  scale_color_manual(values = c("#E69F00" ,"#E69F00" ,"#CC79A7", "#CC79A7")) +
  # makes the boxplots 
  geom_boxplot(width = 0.35, outlier.shape = NA, color = "white", alpha = 0.1, lwd=0.8) +
  annotate("point", x = 1, y = point_estimate[4], colour = "#000000", fill = "#CCCCCC", size = 2, shape = 21) + 
  annotate("point", x = 2, y = point_estimate[3], colour = "#000000", fill = "#CCCCCC", size = 2, shape = 21) +
  annotate("point", x = 3, y = point_estimate[2], colour = "#000000", fill = "#CCCCCC", size = 2, shape = 21) +
  annotate("point", x = 4, y = point_estimate[1], colour = "#000000", fill = "#CCCCCC", size = 2, shape = 21) +
  # plots the point for the mode
  # this is a possible theme of the plot, there are many others
  theme_classic() +
  # changes the labels on the x axis
  scale_y_continuous(limits = c(-0.01 ,0.25),
                     breaks = seq(0, 0.25, 0.05)) +
  scale_x_discrete(labels = c("Family S2" = "Family effect\nreplication study",
                              "Family S1" = "Family effect\noriginal study",
                              "Colony S2" = "Colony effect\nreplication study",
                              "Colony S1" = "Colony effect\noriginal study"),
                   limits = c("Family S2",
                              "Family S1",
                              "Colony S2", 
                              "Colony S1")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("") +
  # label for y axis
  ylab("Explained variation") +
  # flips plot so everything is horizontal
  coord_flip() +
  theme(panel.background = element_rect(colour = "black", size = 1.25, fill = NA),
        axis.text = element_text(colour = "black"),
        legend.position = "none")


MP_effectsize_gg
```


## Bootstrapping Mantel tests for scent isolation by distance approach
```{r , echo=TRUE, message=FALSE, warning=FALSE}

# form geographical distance matrix (in meter)
# rows and cols ordered after levels(as.factor(scent_factors$colony))
geogr_dist_beaches <- matrix(c(0,	1661,	225,	430,	2002,	354,
                               1661,	0,	1634,	1231,	3593,	1636,
                               225,	1634,	0,	454,	1968,	129,
                               430,	1231,	454,	0,	2410,	517,
                               2002,	3593,	1968,	2410,	0,	1957,
                               354,	1636,	129,	517,	1957,	0),
                             nrow = 6, ncol = 6)


# load in data to be transformed to be comparable in a Mantel test with geographical distances
load("RData/objects/pup_colonies_nmds_scaling.RData")

# transfer BeachAge Column from scent_nmds to meta data.frame scent_factors
scent_factors <- cbind(scent_factors, 
                       BeachAge = scent_nmds$BeachAge)
# create index column for meta data frame
scent_factors <- cbind(scent_factors,
                       SampleIndex = 1:length(rownames(scent_factors)))

# iterate/repeat mantel test 
# create data.frame to track mantel results over repeated tests
mantel_btrap_cont <- data.frame(R2 = double(), p = double())

for (iter in 1:999) {
  # sample one individual for each colony and assign index of individual to new vector
  # that scent_Index_mantel_permute can then be used easily to reference indeces in all scent dfs
  scent_Index_mantel_permute <- NULL
  for (i in 1:length(levels(as.factor(scent_factors$colony)))) {
    scent_Index_mantel_permute[i] <- sample(scent_factors$SampleIndex[scent_factors$colony == levels(as.factor(scent_factors$colony))[i]], 
                                            1, 
                                            replace = F)
  } #close i
  
  # create scent profile dissimilarity matrix for 6 differing individuals using Bray-Curtis 
  scent.mantel <- as.matrix(vegdist(scent[scent_Index_mantel_permute,], method = "bray"))
  
  # perform mantel test: scent dissimilarity by metric distance 
  mantel_result <- mantel(scent.mantel, geogr_dist_beaches, 
                          method = "pearson",
                          permutations = 999)
  # store results for rbind in new vector
  mantel_iter_result <- cbind(mantel_result$statistic, mantel_result$signif)
  mantel_btrap_cont <- rbind(mantel_btrap_cont,
                             mantel_iter_result)
} #close iter
colnames(mantel_btrap_cont) <- c("R2", "p-Value")

mantel_btrap_plot <- ggplot(data = mantel_btrap_cont, aes(x = `p-Value`, y = R2)) +
  geom_point(size = 1) + 
  geom_vline(xintercept = 0.05, color = "red", linetype = 3) +
  scale_x_continuous("p-Value",
                     breaks = c(0.05, 0.2, 0.35, 0.5, 0.65, 0.8, 0.95)) +
  theme_classic() +
  theme(panel.background = element_rect(colour = "black", size = 1.25, fill = NA),
        axis.text = element_text(colour = "black"),
        legend.position = "none")
mantel_btrap_plot
```


## Mantel test based on a colony profile approach
```{r Mantel test based on a colony profile approach, echo=TRUE, message=FALSE, warning=FALSE}
## MANTEL TEST BY COLONY LEVEL SCENT PROFILES BASED ON SUMMING INDIVIDUAL PEAKS TO SIMULATE A COLONY PROFILE
## create colony scent profiles for six breeding beaches of Antarctic fur seals

# load in aligned peak data and metadata of pup colonies
load("RData/objects/pup_colonies_nmds_scaling.RData")

# create colony profile by overall peak concentration (or by mean)
fwb_profile <- (apply(scent[which(scent_factors$colony == "FWB"),], 2, function(x) mean(x))) #sum(x[x>0])
johnson_profile <- (apply(scent[which(scent_factors$colony == "johnson"),], 2, function(x) mean(x))) #[x>0]
landing_profile <- (apply(scent[which(scent_factors$colony == "landing_beach"),], 2, function(x) mean(x))) #[x>0]
mainbay_profile <- (apply(scent[which(scent_factors$colony == "main_bay"),], 2, function(x) mean(x))) #[x>0]
natarch_profile <- (apply(scent[which(scent_factors$colony == "natural_arch"),], 2, function(x) mean(x))) #[x>0]
ssb_profile <- (apply(scent[which(scent_factors$colony == "SSB"),], 2, function(x) mean(x))) #[x>0]

colony_profiles <- as.data.frame(rbind(FWB = fwb_profile,
                                       Johnson = johnson_profile,
                                       LandingBeach = landing_profile,
                                       MainBay = mainbay_profile,
                                       NatArch = natarch_profile,
                                       SSB = ssb_profile))
colony_profiles_dist <- as.matrix(vegan::vegdist(colony_profiles, method = "bray"))

# form geographical distance matrix (in meter)
# rows and cols ordered after levels(as.factor(scent_factors$colony))
geogr_dist_beaches <- matrix(c(0,	1661,	225,	430,	2002,	354,
                               1661,	0,	1634,	1231,	3593,	1636,
                               225,	1634,	0,	454,	1968,	129,
                               430,	1231,	454,	0,	2410,	517,
                               2002,	3593,	1968,	2410,	0,	1957,
                               354,	1636,	129,	517,	1957,	0),
                             nrow = 6, ncol = 6)

# Mantel test
mantel(colony_profiles_dist, geogr_dist_beaches, 
       method = "pearson",
       permutations = 999)
```
## Venn diagram for mother-pup pairs
```{r eval=FALSE, include=TRUE}
# Find all peaks in a subset of one colony and write to vector. Repeat for all colonies. Create data.frame that shows overall presence or absence of a peak in the respective colony. Find unique peaks for each colony. -> Venn diagramm
require(nVennR)

load("RData/objects/mom_pup_nmds_scaling.RData")

## convert scent data.frame to logicals. Peaks == 1, No Peaks == 0. Prepending a + will type cast to integer.
scent.log <- + sapply(scent, as.logical)
scent.log <- as.data.frame(scent.log)
rownames(scent.log) <- rownames(scent)

## find all peaks in a subset of one colony
# index with meta data scent_factors
# as function:
colPeaks <- function(colTag){ # as.character(colTag)
  with(scent_nmds, 
       # write output as a vector, apply function to array (list, data.frame...)
       # subset {scent.log} to only include individuals from "SSB" Colony
       as.vector(apply(scent.log[which(BeachAge == as.character(colTag)),], 
                       # apply on each column
                       2, 
                       # apply function with parameter, prepending a
                       # + will type cast to integer (T=1, F=0)
                       function(x) +any(x!= 0)
       )))
}

# create character.vector with colony level names -> transformed to not colony level names but names
# with information about maturity and colony status!
venn_factor_names <- levels(as.factor(scent_nmds$BeachAge))

# create data.frame with unique peaks for each colony using colPeaks function and apply on each element in 
# venn_factor_names vector
venn_factor.peaks <- as.data.frame(sapply(venn_factor_names, colPeaks))

# define rownames:-> peak names
venn_factor.peaks <- cbind(Peaks = colnames(scent.log), venn_factor.peaks)


## Plot Venn Diagramm

# Subset colonies for Venn Diagram according to the example in 'nVennR' Vignette
FWB1 <- subset(venn_factor.peaks, FWB_1 == 1)$Peaks
FWB2 <- subset(venn_factor.peaks, FWB_2 == 1)$Peaks
SSB1 = subset(venn_factor.peaks, SSB_1 == 1)$Peaks
SSB2 <- subset(venn_factor.peaks, SSB_1 == 1)$Peaks


# create Venn Diagram and output as .svg file (vector graphic)
myVenn <- plotVenn(list(FWB1 = FWB1, 
                        FWB2 = FWB2,
                        SSB1 = SSB1, 
                        SSB2 = SSB2
                    
), # close list
nCycles = 9999,
outFile='iter1_mp_ssb_fwb.svg'
) #close plotVenn

# rerun the nVennObj 'myVenn' to increase computation speed and accuracy of the diagramm output
myVenn <- plotVenn(nVennObj = myVenn,
                   outFile = 'RData/figures/iter2_mp_ssb_fwb.svg')

myVenn <- plotVenn(nVennObj = myVenn,
                   outFile = 'RData/figures/iter3_mp_ssb_fwb.svg')

myVenn <- plotVenn(nVennObj = myVenn,
                   labelRegions = F,
                   borderWidth = 2,
                   setColors = c("#D55E00", 
                                 "#E69F00", 
                                 "#0072B2",
                                 "#56B4E9"),
                   outFile = 'RData/figures/Venn_mp_ssb_fwb.svg')
# Iterations don't need to be saved, but can help to understand the process
# Note that the Venn is no object, that can be viewed in R. Instead use a
# program that can handle .svg formats, for example InkScape.
```

## Venn diagram for all pup colonies
```{r eval=FALSE, include=TRUE}
load("RData/objects/pup_colonies_nmds_scaling.RData")

## convert scent data.frame to logicals. Peaks == 1, No Peaks == 0. Prepending a + will type cast to integer.
scent.log <- + sapply(scent, as.logical)
scent.log <- as.data.frame(scent.log)
rownames(scent.log) <- rownames(scent)

## find all peaks in a subset of one colony
# index with meta data scent_factors
# as function:
colPeaks <- function(colTag){ # as.character(colTag)
  with(scent_factors, 
       # write output as a vector, apply function to array (list, data.frame...)
       # subset {scent.log} to only include individuals from "SSB" Colony
       as.vector(apply(scent.log[which(colony == as.character(colTag)),], 
                       # apply on each column
                       2, 
                       # apply function with parameter, prepending a
                       # + will type cast to integer (T=1, F=0)
                       function(x) +any(x!= 0)
       )))
}

# create character.vector with colony level names
colony_names <- levels(as.factor(scent_factors$colony))

# create data.frame with unique peaks for each colony using colPeaks function and apply on each element in 
# colony_names vector
colony.peaks <- as.data.frame(sapply(colony_names, colPeaks))

# define rownames:-> peak names
colony.peaks <- cbind(Peaks = colnames(scent.log), colony.peaks)


## Plot Venn Diagramm

# Subset colonies for Venn Diagram according to the example in 'nVennR' Vignette
FWB <- subset(colony.peaks, FWB == 1)$Peaks
johnson <- subset(colony.peaks, johnson == 1)$Peaks
landing_beach = subset(colony.peaks, landing_beach == 1)$Peaks
main_bay <- subset(colony.peaks, main_bay == 1)$Peaks
natural_arch <- subset(colony.peaks, natural_arch == 1)$Peaks
SSB <- subset(colony.peaks, SSB == 1)$Peaks

# create Venn Diagram and output as .svg file (vector graphic)
myVenn <- plotVenn(list(FWB = FWB, 
                        Johnson = johnson,
                        Landing = landing_beach,
                        MainBay = main_bay,
                        NaturalArch = natural_arch,
                        SSB = SSB
), # close list
nCycles = 9999,
outFile='~/iter1.svg'
) #close plotVenn

# rerun the nVennObj 'myVenn' to increase computation speed and accuracy of the diagramm output
myVenn <- plotVenn(nVennObj = myVenn,
                   outFile = 'RData/figures/iter2.svg')

myVenn <- plotVenn(nVennObj = myVenn,
                   outFile = 'RData/figures/iter3.svg')

myVenn <- plotVenn(nVennObj = myVenn,
                   outFile = 'RData/figures/Venn_allcolonies.svg')
```

## Session information
```{r Session information, echo=FALSE}
sessionInfo()
```

